% Related Work — The Validation Gap (CoRL 2026)
% All citations live-verified via web/arXiv in May 2026, except †classical biostatistics
% refs (Prentice 1989; Buyse & Molenberghs 1998) verified against training knowledge.

\section{Related Work}

\paragraph{Observations of the validation gap.}
The gap between checkpoint-selection by validation loss and peak rollout performance has
been noted independently by several groups. \citet{mandlekar2021} trained imitation
policies on every saved checkpoint and reported that the validation-loss-selected model was
50--100\% worse in success rate than the rollout-best model; the observation appeared in a
single discussion sentence with no systematic follow-up. \citet{zhao2023act} adopted
validation loss as the default checkpoint-selection rule in ACT, acknowledging it as a
working compromise rather than a principled choice. \citet{li2024simpler} compared
validation MSE against rollout success rate for several generalist policies and found low
correlation, using this as the main motivation to build a simulation-based evaluation
environment rather than fix the offline metric. A practitioner synthesis \citep{marchand2024}
confirmed community-side awareness of the problem and the absence of a consensus
alternative. Most recently, \citet{liu2025liberopro} and \citet{lee2025liberoplus}
demonstrated that LIBERO-trained vision-language-action policies scoring above 90\% on the
standard benchmark collapse to near zero under modest perturbations (shifted objects,
instruction paraphrases, changed viewpoints), indirectly showing that the validation-driven
selection pipeline overfits the test distribution. Closest on the empirical side,
\citet{hpil2021} evaluated reward proxies for selecting imitation-learning hyperparameters on
continuous-control benchmarks and found state divergence a useful proxy; that study predates
diffusion and transformer visuomotor policies, treats low-dimensional control rather than
image-based manipulation, and offers no regime characterization, theory link, or pre-registration.
Our work is the first to study this gap systematically across task regimes and modern visuomotor
architectures, compare multiple candidate offline signals organized by environment access, and
frame the question as a calibration problem.

\paragraph{Rollout-based evaluation and efficient alternatives.}
A complementary line of work accepts the cost of rollout evaluation and focuses on making
it tractable. \citet{vincent2024stochastic} provide confidence bounds on a policy's success
probability via Monte Carlo rollouts. \citet{snyder2026binary} extend this to a
sample-efficient sequential testing framework based on safe anytime-valid inference,
reducing evaluation burden by up to 70\%. \citet{patil2026golden} search over the initial
noise distribution of a frozen Diffusion Policy via Monte Carlo evaluation, improving
success on 38 of 43 tasks. \citet{wang2026ppguide} train a performance predictor on
rollout-labeled action chunks and use it to steer policy execution at inference time.
\citet{anwar2025} allocate a fixed rollout budget across tasks using active experiment
selection. A parallel real-to-sim line builds simulators whose success correlates with real-world
performance: SIMPLER~\citep{li2024simpler}, PolaRiS~\citep{polaris2025}, and high-fidelity
sim-to-real benchmarking~\citep{nvidia2025eval} all validate that simulated rankings track real
ones, and RoboArena crowdsources real-robot evaluation directly. Our question is upstream of all of
these: not how to evaluate cheaply by rollout or sim, but whether any signal computable
\emph{without} rolling out at all suffices. These works collectively establish that
rollout-based and sim-based evaluation are the trusted standards and are becoming cheap enough to
use routinely. Our contribution is the orthogonal
question: \emph{can any signal computable without rollouts recover the deployment outcome?}
If the answer is yes, training pipelines become substantially cheaper; if no, this cluster's
rollout-based approach is the principled one, and we provide the calibrated evidence for
that conclusion.

\paragraph{Offline and behavioral evaluation metrics.}
\citet{tiezzi2025neme} proposed a learned offline meta-evaluator for imitation policies in
humanoid human-robot interaction, explicitly motivated by validation loss's inadequacy. Their
approach scores policies using Dynamic Time Warping of joint trajectories and is the closest
prior work to our M5/M7 metrics. It addresses a single task domain, proposes one metric, and
does not compare against other offline signals. \citet{wang2025roboeval} introduced
RoboEval, a structured evaluation framework that augments binary success with behavioral
metrics --- coordination quality, trajectory efficiency, and spatial precision --- and
showed they correlate with success in 59.4\% of task-metric pairs across bimanual
manipulation. Critically, RoboEval's metrics are computed \emph{during} closed-loop
execution: they require running the policy in the environment. We study the complementary
case: metrics computable purely from the policy and held-out demonstrations, with no
environment interaction whatsoever.

\paragraph{Compounding-error theory.}
That offline action error compounds into closed-loop failure is classical: \citet{ross2010efficient,ross2011reduction}
established the quadratic-in-horizon return gap, \citet{spencer2021feedback} characterized the three
regimes of covariate shift, and \citet{simchowitz2025pitfalls} showed exponential compounding is
unavoidable under unstable continuous dynamics. Closest to us, \citet{tan2026gain} prove that
horizon-$H$ failure probability factorizes into a gain-dependent amplification term and the validation
loss, formalizing why validation loss cannot rank checkpoints. We take this theory as given. Our
contribution is orthogonal and empirical: a systematic calibration of eight offline metrics against
ground-truth success across two modern visuomotor architectures and eight task regimes, and a
rollout-free predictor that operationalizes the theory's prediction.

\paragraph{Runtime failure detection.}
A parallel line of work applies OOD signals at execution time rather than at checkpoint-selection time. FAIL-Detect \citep{faildetect2025} frames runtime failure detection as sequential OOD detection, distilling policy inputs and outputs into scalar signals that correlate with failures and using conformal prediction for statistical guarantees. Subsequent work extends this to generative policies~\citep{failpred2025}. Both detect coherence failures while a policy is executing. Our M4 (latent Mahalanobis distance) applies the same OOD principle one stage earlier --- at training time, to select which checkpoint to execute in the first place.

\paragraph{Policy architecture and benchmarks.}
This work trains and evaluates Diffusion Policy \citep{chi2023diffusion}, the modal
architecture for visuomotor manipulation, on the LIBERO benchmark \citep{liu2023libero} and
the Robomimic benchmark \citep{mandlekar2021}. Architecture and benchmark choices are
frozen before training; the scientific question is about the offline-to-online mapping, not
about architecture comparisons.

\paragraph{Surrogate-endpoint calibration.}
The framing we import from drug development defines validation loss as a
\emph{surrogate endpoint} and rollout success rate as the \emph{true endpoint}.
\citet{prentice1989} established the formal criteria for when a surrogate is acceptable.
\citet{buyse1998} developed the meta-analytic framework for surrogate validation, defining
both individual-level and trial-level association requirements that we operationalise through
our per-run Spearman correlations (H3) and cross-task composite predictor (H4).
