% Mandatory statements (CoRL / general venue requirements).

\paragraph{Reproducibility and code availability.} The full pipeline --- data adapters, both
policy architectures, the eight offline metrics, the rollout harness, and the pre-registered
analysis --- is released as an open, pip-installable package with a preflight readiness gate
and a synthetic CPU path that reproduces every code path without a GPU. The analysis script and
the configuration that fixes all pre-registered choices are committed and SHA-256-hashed before
training begins; the hash is recorded in the project log. Each released artifact pairs every
saved checkpoint with its offline metrics and its twenty-rollout success outcome, so a reader
can evaluate a new offline metric against ground-truth success without retraining.

\paragraph{Data availability.} All tasks use public simulation benchmarks (LIBERO and
Robomimic) with publicly available demonstrations. No new datasets are collected. The external
real-robot validation uses already-published policy checkpoints and success rates; we
redistribute none of that data, only pointers to the original releases.

\paragraph{Ethics.} The study is simulation-only and involves no human subjects, no personal
data, and no physical robot deployment. We see no dual-use concern: the work makes policy
evaluation more honest, not more capable. The main societal risk is misplaced trust in an
offline metric, which the calibration framework is designed to prevent rather than create.

\paragraph{AI-usage disclosure.} AI-assisted tools (a large language model) were used to help
implement the codebase, conduct and verify the literature search, and draft and edit the
manuscript. Every cited reference was verified against its primary source; the authors take
full responsibility for all claims, the experimental design, and the final text. No
experimental result reported here was produced by an AI tool.

\paragraph{Author contributions (CRediT).} E.K.: conceptualization, methodology, software,
formal analysis, writing --- original draft, writing --- review and editing.

\paragraph{Funding.} This work received no external funding and used consumer-grade compute.

\paragraph{Conflict of interest.} The authors declare no competing interests.
