% HONEST framing: the bound is standard. We recall it, attribute it fully, and use it to
% state the calibration prediction our experiments test. We do NOT claim the bound as novel.

The validation gap is not a new phenomenon in need of a new theory; it is an instance of the
compounding-error effect studied since the origins of imitation learning. We recall the relevant
bound, attribute it, and draw out the prediction it makes for \emph{offline metrics} --- the
prediction our experiments are built to test. The contribution of this paper is empirical, not
theoretical.

\paragraph{Setup.} Let the dynamics be $s_{t+1}=f(s_t,a_t)$ with $f$ Lipschitz,
$\|f(s,a)-f(s',a')\|\le L_s\|s-s'\|+L_a\|a-a'\|$, let $\pi^\star$ be the expert,
$g(s):=f(s,\pi^\star(s))$ the expert closed-loop map with $L:=\mathrm{Lip}(g)$, and let a learned
policy have per-step action error $\varepsilon$ along its rollout. With $\delta_t$ the state
deviation from the expert rollout and $\delta_0=0$, the standard error-propagation recurrence
$\delta_{t+1}\le L\,\delta_t+L_a\varepsilon$ unrolls to
\[
\delta_H \;\le\; L_a\,\varepsilon\sum_{k=0}^{H-1}L^{k}
\;=\;\begin{cases} L_a\,\varepsilon\,(L^{H}-1)/(L-1), & L\neq 1,\\[0.6ex] L_a\,\varepsilon\,H, & L=1,\end{cases}
\tag{1}\label{eq:dev}
\]
which exhibits three regimes: bounded deviation for self-correcting dynamics ($L<1$), linear growth
at the stability boundary ($L=1$), and exponential growth for expansive, contact-rich dynamics
($L>1$). This is classical. The recurrence is the discrete Gr\"onwall / error-propagation argument;
the quadratic-in-horizon return version is due to \citet{ross2010efficient,ross2011reduction}; the
three-regime characterization of covariate shift is \citet{spencer2021feedback}; and the
unavoidability of exponential compounding under unstable continuous dynamics is
\citet{simchowitz2025pitfalls}. Most directly, \citet{tan2026gain} show that the probability of
horizon-$H$ task failure factorizes into a gain-dependent amplification term and the validation
loss, making precise that validation loss alone cannot rank checkpoints. We claim none of this as
new.

\paragraph{What it predicts for offline metrics.} The value of \eqref{eq:dev} here is as a lens on
\emph{which offline quantity to measure}. Validation loss estimates $\varepsilon$, the per-step
error --- the numerator of \eqref{eq:dev}. Deployment success is governed by whether the
accumulated deviation $\delta_t$ stays inside the task tolerance --- the left-hand side. The two
coincide only when $L\le 1$ and $H$ is small, so two predictions follow that we test directly:

\begin{itemize}\itemsep2pt
\item[(P1)] The gap should grow with horizon and contact along the amplification factor
$(L^H-1)/(L-1)$. This is the regime dependence H2 evaluates; the theory fixes its sign and shape in
advance rather than leaving it a hunch.
\item[(P2)] A metric that estimates $\delta$ rather than $\varepsilon$ should predict success where
validation loss fails. Open-loop replay distance (M5) computes accumulated deviation under exactly
the recurrence above, and latent drift (M4) measures departure from the training manifold, the event
large $\delta_t$ triggers. This is the coherence hypothesis, now with a mechanism: coherence metrics
estimate the integral, local metrics the integrand.
\end{itemize}

One honest caveat strengthens the case: validation loss estimates $\varepsilon$ on the expert's
state distribution, while \eqref{eq:dev} uses the on-policy error, typically larger under covariate
shift --- so \eqref{eq:dev} is, if anything, optimistic about validation loss.

\paragraph{An identifiability limit, and the line it draws.}
The bound above is standard. The observation we build on is not the bound but a partition of the
candidate metrics by whether they \emph{query the environment}. Call a metric \emph{rollout-free}
if it is a functional of the policy evaluated only on the expert-demonstration distribution
$d^\star$; M1--M4, M6, and M8 are rollout-free, whereas M5 and M7 execute the policy in the
simulator. This distinction has a sharp consequence.

\begin{proposition}[Rollout-free metrics cannot rank in the expansive regime]\label{prop:imposs}
Fix a task with expansiveness $L>1$ and horizon $H$. For every resolution $\zeta>0$ there exist two
policies that are $\zeta$-indistinguishable under \emph{every} rollout-free metric --- they agree to
within $\zeta$ on $d^\star$ and so produce identical metric values up to $\zeta$ --- yet whose
closed-loop success rates differ by a constant independent of $\zeta$. No rollout-free metric can
therefore consistently rank such checkpoints by success. A metric that queries the environment, even
open-loop, is not subject to this limit.
\end{proposition}

\begin{proof}[Proof sketch]
Let the two policies coincide with $\pi^\star$ on the support of $d^\star$ up to a perturbation of
size $\zeta$, and differ only off $d^\star$: one steers toward a failure set, the other
self-stabilizes. Because $L>1$, the $\zeta$ on-manifold difference is amplified and the closed loop
leaves any fixed neighborhood of the expert manifold in $O(\log(1/\zeta)/\log L)$ steps
($\le H$ for $\zeta$ small), after which the off-manifold behavior --- unconstrained by $d^\star$ ---
fixes the outcome and the success rates separate by $\Omega(1)$. A rollout-free metric sees only the
$\zeta$ difference on $d^\star$; an environment-querying metric observes the off-manifold divergence.
\end{proof}

This is a selection-theoretic reading of the long-known insufficiency of expert-distribution data
\citep{ross2011reduction,simchowitz2025pitfalls}; what we take from it is the \emph{line it draws
through our metric set}. It predicts that in contact-rich, long-horizon tasks ($L>1$, large $H$) no
rollout-free metric --- validation loss included --- can be relied on, while M5 and M7, which touch
the simulator, can. It is the formal version of this paper's ``horizon wall'' outcome, and it makes
H3's metric-by-regime comparison a test of a stated boundary rather than an open fishing expedition.
To our knowledge no prior work empirically calibrates these offline metrics against ground-truth
success across modern visuomotor policies and task regimes, nor organizes them by this boundary,
which is the gap this paper fills.
